On Misuses of the Kolmogorov--Smirnov Test for One-Sample Goodness-of-Fit

### 2023-04-08 manuscript/manuscript.tex
cleaned up desired plots

The rest of the paper is organized as follows. Section~\ref{sec:fitted}
investigates the scenario where the hypothesized distribution has unspecified
parameters. Both parametric and nonparametric bootstrap are available to fix the
issue. Section~\ref{sec:dependence} investigates the scenario where the data of
the empirical distribution are serially dependent. An bootstrap procedure
employing a working ARMA model to account for dependence is proposed as a
working solution. Section~\ref{sec:fittedwithdependence}
explores the case where a combination of the first two scenarios occurs.
Section~\ref{sec:conclusion} concludes with a discussion.

@@

The consequence of using the wrong null distribution for the KS statistic $D_n$
in~\eqref{eq:ks_fitted} can be illustrated through a simple simulation study. A
random sample $X_1, \ldots, X_n$ was generated from the standard normal
distribution with sample size $n = 100$. The sample mean and sample standard
deviation were used as $\hat\theta_n$. The p-value of the test
statistic~$D_n$ in~\eqref{eq:ks_fitted} was obtained by naively calling the
\texttt{ks.test()} function in~R with the fitted normal distribution as the
hypothesized distribution. That is, the hypothesized distribution was
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample
variance. This experiment was repeated 1000 times and the histogram of the
1000 p-values are displayed in Figure~\ref{fig:hist_fitted}. If the test were
valid, the p-values would be uniformly distributed in the unit
interval. Clearly, the pattern of the 1000 p-values are very different from what
one would expect to see from 1000 draws from the standard uniform
distribution.

@@

Nonparametric bootstrap can also be used to approximate the null distribution
of the test statistic with the bias correction from \citet{babu2004goodness}:
\begin{enumerate}
\item
  Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
  with replacement
\item
  Fit $F_\theta$ to $X_1^*,...,X_n^*$ and obtain estimate 
	$\hat\theta_n^*$ of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of
  $X_1^*, \ldots,  X_n^*$. \eds{verify edit} \jy{Good.}
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x | \sqrt{n}\left(F_n^* (x)- F_{\hat\theta_n^*}(x)\right) - B_n(x) |.
  \]
  where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known
  bias term \citep{babu2004goodness}.
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of $D_n^*$ to approximate the null distribution of the observed
  statistic. 
\end{enumerate}
The p-value of $D_n$ is, again, approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$. Note that the step~1 and
step~4 are different from the parametric bootstrap version.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_fitted}
  \caption{Histograms of p-values for the KS test with unspecified parameters in
    the hypothesized distribution based on 1000 replicates of sample size
    $n = 100$. Naive (left): KS test with KS statistic calculated based on fitted
    parameters. Parametric bootstrap (center): KS test with parametric
    bootstrap. Nonparametric bootstrap (right): KS test with nonparametric
    bootstrap using the bias correction from \citet{babu2004goodness}.
    \eds{perhaps use density hist vs
	freq. hist (here and elsewhere)?}}
  \label{fig:hist_fitted}
\end{figure}

\jy{Anthony, could you label the panels as ``Parametric bootstrap'' and
  ``Nonparametric bootstrap'' for the center and right panels?}

\jy{Simulation study: let's use two distributions:
  $N(8, 8)$ and $\Gamma(8, 1)$. These two distributions have the same means and
  variances. So we could use them to study powers of the tests too.}

In the same simulation study, p-values were obtained for the 1000 replicates
with both the parametric and nonparametric bootstrap procedures. The center and
right panels of Figure~\ref{fig:hist_fitted} display the histograms of the 1000
p-values from the parametric and nonparametric bootstrap procedures,
respectively. This time, both samples of p-values appear to come from
standard uniform distributions.

@@

\jy{For the same two marginal distributions $N(8, 8)$ and $\Gamma(8, 1)$, let's
  generate data from dependence structures of AR(1) with two coefficients
  $\psi \in \{ -0. 5, 0.5\}$.}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_correlation}
  \caption{Histograms of p-values for the KS test applied to data generated from
    AR(1) processes with AR coefficients
    $\psi \in \{-0.3, -0.2, -0.1, 0, 0.1, 0.2, 0.3\}$
    based on 1000 replicates of sample size $n = 100$.}
  \label{fig:hist_correlation}
\end{figure}

@@

\jy{Here we work on the same marginal distributions $N(8, 8)$ and $\Gamma(8, 1)$
  with different dependence structures:
  ARMA(1, 1); MA(2); AR(2)
  }
Continuing with the simulation study, 
Throughout the simulation we assume a working AR(1) dependence structure
regardless of the true dependence. This is demonstrated with different dependence
structures in Figure~\ref{fig:hist_ma1_arma_ar2_D}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_ar1_D}
  \caption{The Naive plot (left) is the histogram of p-values from the
  standard KS test. The Copula plot (right) is the histogram of p-values from
  implementing the parametric \eds{semiparametric?} bootstrap with copulas for
	dependence. In each case,
  $1000$ replicate tests were performed with the data generated from an AR(1)
  process of the standard normal distribution with $\psi = .5$. The sample size
  is $n = 100$ and $B = 1000$ bootstrap samples are obtained for each test.}
  \label{fig:hist_ar1_D}
\end{figure}

Figure~\ref{fig:hist_ar1_D} displays the results of the copula remedy for
dependent data. The data is generated from the standard normal distribution with
an AR(1) dependence structure where $\psi = .5$. The copula used is to model
dependence is the normal copula. The distribution of p-values is again expected
to be $U(0, 1)$ for a valid test. The Naive plot is clearly not uniform and
reinforces the results shown in Figure~\ref{fig:hist_correlation}. The Copula
plot, which implements the aforementioned procedure to correct for dependence
in the data, appears to be uniform and restores the size of the KS test.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_ma1_arma_ar2_D}
  \caption{The Naive plots are the histograms of p-values from the standard KS
  test. The Copula plots are the histograms of p-values from parametric
  bootstrap with copulas for dependence. The data is generated from an MA(1)
  process (left) with $\theta = .5$, ARMA(1, 1) process (middle) with $\psi =
  .5$ and $\theta = .3$, and AR(2) process (right) with $\psi = (.5, .3)$ of the
  standard normal distribution. In each case, $1000$ replicate tests were
  performed with sample size $n = 100$ and $B = 1000$ bootstrap samples.}
  \label{fig:hist_ma1_arma_ar2_D}
\end{figure}

\jy{to be rewritten when the figures are generated.}
\jy{I suggest that we keep using gamma as the marginal distribution
  throughout. Elizabeth, let's finalize the settings next time we meet. I've
  implemented a concise version of the test.}

The procedure detailed in this section can also provide a reasonable
approximation in cases where the true dependence structure is not far from
AR(1). Figure~\ref{fig:hist_ma1_arma_ar2_D} shows the distribution of p-values
for dependence structures MA(1), ARMA(1, 1), and AR(2). As expected,
naively performing the KS test without correcting for dependence provides plots
of p-values that deviate from the uniform distribution. In the case of MA(1)
and ARMA(1, 1), the true dependence structure is close enough to our
assumption of AR(1) that the bootstrap procedure provides a reasonable
approximation. However, the AR(2) copula plot shows the limitation of this
technique as no AR(1) process can approximate an AR(2) process unless the
second-order coefficient is close to zero.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_gamma_gev_D}
  \caption{The Naive plots are the histograms of p-values from the standard KS
  test. The Copula plots are the histograms of p-values from parametric
  bootstrap with copulas for dependence. The data is generated from $Gamma(3,
  1)$ (left) and $GEV(0, .2, 1)$ (right) with a correlation coefficient of $r =
  0.5$. In each case, $1000$ replicate tests were performed with sample size $n
  = 100$ and $B = 1000$ bootstrap samples.}
  \label{fig:hist_gamma_gev_D}
\end{figure}

To further demonstrate the effectiveness of the procedure, we perform similar
simulations on other families of distributions.
Figure~\ref{fig:hist_gamma_gev_D} displays the results of tests using dependent
data generated from the gamma and the generalized extreme value distributions.
The procedure corrects the distribution of p-values for both distributions and
is shown to be applicable regardless of the family of distribution tested.

@@

\jy{For the simulation study, repeat the study at the end of the last section.}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_ar1_FD}
  \caption{The Super Naive plot (left) is the histogram of p-values from the
  standard KS test with fitted parameters. The Naive Parametric plot (middle) is
  the histogram of p-values from implementing parametric bootstrap. The Copula
  plot is the histogram of p-values from implementing parametric bootstrap with
  copulas for dependence and correcting for fitted parameters. In each case,
  $1000$ replicate tests were performed with the data generated from an AR(1)
  process of the standard normal distribution with $\psi = .5$. The sample size
  is $n = 100$ and $B = 1000$ bootstrap samples are obtained for each test.}
  \label{fig:hist_ar1_FD}
\end{figure}

Figure~\ref{fig:hist_ar1_FD} shows the results of our simulation on data
simulated from an AR(1) process. As should be expected, naively fitting
parameters while providing no adjustment for dependent data invalidates the KS
test. As well, only using the parametric bootstrap remedy for fitted parameters
from Section~\ref{sec:fitted} seems to provide weaker results than the
copula remedy presented above. The results show that copula remedy generates
uniform p-values and restores the size of the KS test.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_ma1_arma_ar2_FD}
  \caption{The Super Naive plots (left) are the histograms of p-values from the
  standard KS test with fitted parameters. The Naive Parametric plots (middle)
  are the histograms of p-values from implementing parametric bootstrap. The
  Copula plots (bottom) are the histograms of p-values from implementing
  parametric bootstrap with copulas for dependence and correcting for fitted
  parameters. The data is generated from an MA(1) process (left) with
  $\theta = .5$, ARMA(1, 1) process (middle) with $\psi = .5$ and $\theta = .3$,
  and AR(2) process (right) with $\psi = (.5, .3)$ of the standard normal
  distribution. In each case, $1000$ replicate tests were performed with sample
  size $n = 100$ and $B = 1000$ bootstrap samples.}
  \label{fig:hist_ma1_arma_ar2_FD}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_gamma_gev_FD}
  \caption{The Super Naive plots (left) are the histograms of p-values from the
  standard KS test with fitted parameters. The Naive Parametric plots (middle)
  are the histograms of p-values from implementing parametric bootstrap. The
  Copula plots are the histograms of p-values from implementing parametric
  bootstrap with copulas for dependence and correcting for fitted parameters.
  The data is generated from $Gamma(3, 1)$ (left) and $GEV(0, .2, 1)$ (right)
  with a correlation coefficient of $r = 0.5$. In each case, $1000$ replicate
  tests were performed with sample size $n = 100$ and $B = 1000$ bootstrap
  samples.}
  \label{fig:hist_gamma_gev_FD}
\end{figure}

It is also possible to apply our procedure to other families of distributions.
The data in Figure~\ref{fig:hist_gamma_gev_FD} is generated from the gamma and
generalized extreme value distributions. The results are favorable and show that
the procedure outlined as a remedy for fitted parameters and spatially dependent
can be adapted for various families of distributions.

@@

>>>>

The rest of the paper is organized as follows. Section~\ref{sec:fitted}
investigates the scenario where the hypothesized distribution has unspecified
parameters. Both parametric and nonparametric bootstrap are available to fix the
issue. Section~\ref{sec:dependence} investigates the scenario where the data of
the empirical distribution are serially dependent. A bootstrap procedure
employing a working ARMA model to account for dependence is proposed as a
working solution. Section~\ref{sec:fittedwithdependence}
explores the case where a combination of the first two scenarios occurs.
Section~\ref{sec:conclusion} concludes with a discussion.

@@

The consequence of using the wrong null distribution for the KS statistic $D_n$
in~\eqref{eq:ks_fitted} can be illustrated through a simple simulation study. A
random sample $X_1, \ldots, X_n$ was generated from the standard normal
distribution with sample size $n = 100$. The sample mean and sample standard
deviation were used as $\hat\theta_n$. The p-value of the test
statistic~$D_n$ in~\eqref{eq:ks_fitted} was obtained by naively calling the
\texttt{ks.test()} function in~R with the fitted normal distribution as the
hypothesized distribution. That is, the hypothesized distribution was
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample
variance. This experiment was repeated 1000 times and the histogram of the
1000 p-values are displayed in Figure~\ref{fig:pp_f}. If the test were
valid, the p-values would be uniformly distributed in the unit
interval. Clearly, the pattern of the 1000 p-values are very different from what
one would expect to see from 1000 draws from the standard uniform
distribution.

@@

Nonparametric bootstrap can also be used to approximate the null distribution
of the test statistic except that a bias correction is needed
\citet{babu2004goodness}. Details of the nonparametric bootstrap procedure is
in the Appendix. The results from nonparametric bootstrap are similar to those
from the parametric bootstrap and, hence, are omitted.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_fitted}
  \caption{PP-plot for the p-values obtained from KS tests with unspecified
    parameters with sample size $n = 200$ based on 1000 replicats. The data
    generating models were $N(8, 8)$ and $\Gamma(8, 1)$.
  }
  \label{fig:pp_f}
\end{figure}

In the same simulation study, p-values were obtained for the 1000 replicates
with both the parametric and nonparametric bootstrap procedures. The center and
right panels of Figure~\ref{fig:pp_f} display the histograms of the 1000
p-values from the parametric and nonparametric bootstrap procedures,
respectively. This time, both samples of p-values appear to come from
standard uniform distributions.

@@

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_correlation}
  \caption{PP-plots for the p-values obtained from KS test with serial
    dependence discarded and sample size $n = 200$ based on 1000
    replicates. The data have marginal distributions $N(8, 8)$ and
    $\Gamma(8, 1)$, with an AR(1) dependence structure on the standard normal
    scale. The AR coefficients were $\psi \in \{-0.3, -0.1,  0.1,  0.3\}$.
  }
  \label{fig:pp_s}
\end{figure}

@@

\jy{describe the simulation setting here.}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_correlation}
  \caption{PP-plots for the p-values obtained from KS test with serial
    dependence accounted for through a working ARMA model and $n = 200$ based on
    1000 replicates. The data have marginal distributions $N(8, 8)$ and
    $\Gamma(8, 1)$. The dependence structure on the standard normal scale was
    characterized by ARMA(1, 1) model, AR(2) model, and MA(1) model.
  }
  \label{fig:pp_ss}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_correlation}
  \caption{PP-plots for the p-values obtained from KS test with unspecified
    parameters and serial
    dependence accounted for through a working ARMA model and $n = 200$ based on
    1000 replicates. The data have marginal distributions $N(8, 8)$ and
    $\Gamma(8, 1)$. The dependence structure on the standard normal scale was
    characterized by ARMA(1, 1) model, AR(2) model, and MA(1) model.
  }
  \label{fig:pp_ss}
\end{figure}

@@

@@

\appendix

\section{Nonparametric Bootstrap for KS Test with Fitted Parameters}

Using the same notations in the text, the procedure is summarized as follows
\citep{babu2004goodness}.
\begin{enumerate}
\item
  Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
  with replacement
\item
  Fit $F_\theta$ to $X_1^*,...,X_n^*$ and obtain estimate 
	$\hat\theta_n^*$ of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of
  $X_1^*, \ldots,  X_n^*$. 
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x | \sqrt{n}\left(F_n^* (x)- F_{\hat\theta_n^*}(x)\right) - B_n(x) |.
  \]
  where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known
  bias term \citep{babu2004goodness}.
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of $D_n^*$ to approximate the null distribution of the observed
  statistic. 
\end{enumerate}
The p-value of $D_n$ is, again, approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$. Note that the step~1 and
step~4 are different from the parametric bootstrap version.
